manuscript_figures

### 2024-07-10 scripts/manuscript_figures.Rmd
updated mmanuscript figures and habitat productivity index

#load in morph data and select for SUGK
morph.filt <- morph %>%
        filter(SP_CODE == "SL")

@@

# merge morph filt and quad.filt to calculate biomass
merge.morph.quad <- left_join(morph.filt, quad.filt)

@@

# Identify the outliers
outliers <- c(45, 37, 29,37)

@@

# Extract the p-value for the kelp biomass coefficient
p_value_kelp_biomass <- lmm_summary$coefficients["log(mean_kelp_biomass + 1.1)", "Pr(>|t|)"]

@@

  # p-value
  annotate("text", x = Inf, y = -Inf, 
           label = paste0("p = ", format(p_value_kelp_biomass, scientific = FALSE, digits = 2)), 
           hjust = 1.3, vjust = -39, size = 4) +

@@

# Time series plot for kelp biomass with standard error bars from 2016-2023
kelp.timeseries.2016 <- ggplot(aggregated.kelp.morph, aes(x = YEAR, y = mean_kelp_biomass, shape = SITE)) +
  geom_line(aes(group = SITE), color = "black") +
  geom_point(size = 6) +
  geom_errorbar(aes(ymin = mean_kelp_biomass - se_kelp_biomass, 
                    ymax = mean_kelp_biomass + se_kelp_biomass), 
                width = 0.2) +
  scale_shape_manual(values = shape_values) +
  labs(x = "Year", y = expression("Mean Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
  theme_bw() +
  scale_x_continuous(breaks = unique(aggregated.kelp.morph$YEAR)) +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.position = c(1, 1),  # Positioning legend inside the plot area at the top right
    legend.justification = c(1, 1),  # Anchoring the legend to the top right corner
    legend.background = element_rect(fill = "white", color = "black")  # Optional: adding a background for better visibility
  );kelp.timeseries.2016

@@

# Time series plot for kelp biomass with standard error bars for 2019-2023
kelp.timeseries.2019_2023 <- ggplot(filtered_data, aes(x = YEAR, y = log_mean_kelp_biomass, shape = SITE)) +
  geom_line(aes(group = SITE), color = "black") +
  geom_point(size = 8) +
  geom_errorbar(aes(ymin = log_mean_kelp_biomass - log_se_kelp_biomass, 
                    ymax = log_mean_kelp_biomass + log_se_kelp_biomass), 
                width = 0.2) +
  scale_shape_manual(values = shape_values) +
  labs(x = "Year", y = expression("Log Mean Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18)
  );kelp.timeseries.2019_2023

@@

# Plotting the mean rates of change with error bars (standard error horizontally)
rate_of_change <- ggplot(slope_results, aes(x = rate_of_change_per_year, y = mean_nonlog_avg_kelp_biomass, shape = SITE)) +
  geom_point(size = 8) +
  geom_errorbarh(aes(xmin = rate_of_change_per_year - se_rate_of_change, xmax = rate_of_change_per_year + se_rate_of_change), height = 0.9) +
  labs(
    x = expression("Rate of Change ("~yr^-1 ~")"),
    y = expression("Mean Kelp Biomass (g/"~m^2~")"), shape = "Site") +
  theme_bw() +
  scale_shape_manual(values = shape_values, guide = guide_legend(title = "Site")) +   # Use shape values and customize legend
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18), ,
    legend.position = "none");rate_of_change

@@

# Organize rugosity data and make linear regression figure for fish biomass and rugosity
```{r}
# Calculate mean rugosity complexity scores across depths for each transect, grouped by site and year
mean_rugosity <- rugosity %>%
  group_by(SITE, YEAR, TRANSECT) %>%
  summarize(mean_rugosity = mean(c(X0M, X4M, X8M, X12M, X16M, X20M, X24M, X28M, X32M, X36M, X40M), na.rm = TRUE))

# change transect from integer to chr
mean_rugosity$TRANSECT <- as.character(mean_rugosity$TRANSECT)

# Left jon fish and rugosity data
merged.dat.rug <- left_join(mean_rugosity, fish.filt)

# Make figure 4 and run statistical analysis
```{r}
# Aggregate the data by year, site, and transect and compute mean values
aggregated.data.rug <- merged.dat.rug %>%
  group_by(YEAR, SITE, TRANSECT) %>%
  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
            mean_rugosity = mean(mean_rugosity, na.rm = TRUE)) %>%
  ungroup()

# Fit the linear mixed effects model
lmm_model <- lmer(mean_fish_biomass ~ mean_rugosity + SITE + (1 | SITE/YEAR), data = aggregated.data.rug)
#check normality
qqPlot(residuals(lmm_model)) #data is normal no need to transform

# Fit the linear mixed effects model with log transformation
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_rugosity + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated.data.rug)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal

# Identify the outliers
outliers <- c(25, 37, 24, 35)

# remove outliers
data_clean.rug <- aggregated.data.rug[-outliers, ]

# rerun model with log and without outliers
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_rugosity + 1.1) + SITE + (1 | SITE/YEAR), data = data_clean.rug)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal

# Extract the p-value for the rugosity
p_value_rugosity <- lmm_summary$coefficients["log(mean_rugosity + 1.1)", "Pr(>|t|)"]

#create scatter plot
figure.4 <- ggplot(aggregated.data.rug, aes(x = log(mean_rugosity + 1.1), y = log(mean_fish_biomass + 1.1))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +

  # p-value
  annotate("text", x = Inf, y = -Inf, 
           label = paste0("p = ", format(p_value_rugosity, scientific = FALSE, digits = 2)), 
           hjust = 1.3, vjust = -38.5, size = 4) +

  labs(x = "Rugosity Complexity", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Rugosity Complexity") +
  theme_classic() +
  theme(plot.margin = margin(t = 10, r = 30, b = 10, l = 10, unit = "pt")) +
  coord_cartesian(clip = 'off');figure.4

# Save the plot to a file
ggsave("output/figures/figure_4.png", plot = figure.4, width = 8, height = 6)
```

# Filter quad data for all inverts and use to correlate to fish biomass
```{r}
#invert.filt <-

@@

# Plot two plots together using patchwork for production and HPI
```{r}
# Combine the plots with patchwork for all years of kelp and rate of change
production.hpi.nonlog <- production_plot | hpi.plot.nonlog +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi.nonlog

>>>>

#load in morph data and select for SUGK
morph.filt <- morph %>%
        filter(SP_CODE == "SL" | SP_CODE == "LADI")

@@

# merge morph filt and quad.filt to calculate biomass
merge.morph.quad <- rbind.fill(morph.filt, quad.filt)

@@

# Identify the outliers
outliers <- c(48,19)

@@

# Extract the p-value for the kelp biomass coefficient
p_value_kelp_biomass <- ""

@@

  # p-value
  annotate("text", x = Inf, y = -Inf, 
           label = paste0("p < 0.0001", format(p_value_kelp_biomass, scientific = FALSE, digits = 2)), 
           hjust = 1.3, vjust = -39, size = 4) +

@@

# Time series plot for kelp biomass with standard error bars from 2016-2023
kelp.timeseries.2016 <- ggplot(aggregated.kelp.morph, aes(x = YEAR, y = mean_kelp_biomass, shape = SITE)) +
  geom_line(aes(group = SITE), color = "black") +
  geom_point(size = 6) +
  geom_errorbar(aes(ymin = mean_kelp_biomass - se_kelp_biomass, 
                    ymax = mean_kelp_biomass + se_kelp_biomass), 
                width = 0.2) +
  scale_shape_manual(values = shape_values) +
  labs(x = "Year", y = expression("Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
  theme_bw() +
  scale_x_continuous(breaks = unique(aggregated.kelp.morph$YEAR)) +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.position = c(1, 1),  # Positioning legend inside the plot area at the top right
    legend.justification = c(1, 1),  # Anchoring the legend to the top right corner
    legend.background = element_rect(fill = "white", color = "black")  # Optional: adding a background for better visibility
  );kelp.timeseries.2016

@@

# Time series plot for kelp biomass with standard error bars for 2019-2023
kelp.timeseries.2019_2023 <- ggplot(filtered_data, aes(x = YEAR, y = log_mean_kelp_biomass, shape = SITE)) +
  geom_line(aes(group = SITE), color = "black") +
  geom_point(size = 8) +
  geom_errorbar(aes(ymin = log_mean_kelp_biomass - log_se_kelp_biomass, 
                    ymax = log_mean_kelp_biomass + log_se_kelp_biomass), 
                width = 0.2) +
  scale_shape_manual(values = shape_values) +
  labs(x = "Year", y = expression("Kelp Biomass (g/" ~ m^2 ~ ")"), shape = "Site") +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18)
  );kelp.timeseries.2019_2023

@@

# Plotting the mean rates of change with error bars (standard error horizontally)
rate_of_change <- ggplot(slope_results, aes(x = rate_of_change_per_year, y = mean_nonlog_avg_kelp_biomass, shape = SITE)) +
  geom_point(size = 8) +
  geom_errorbarh(aes(xmin = rate_of_change_per_year - se_rate_of_change, xmax = rate_of_change_per_year + se_rate_of_change), height = 0.9) +
  labs(
    x = expression("Rate of Change ("~yr^-1 ~")"),
    y = expression("Kelp Biomass (g/"~m^2~")"), shape = "Site") +
  theme_bw() +
  scale_shape_manual(values = shape_values, guide = guide_legend(title = "Site")) +   # Use shape values and customize legend
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18), ,
    legend.position = "none") +
  scale_x_continuous(limits = c(-0.5, 0.0), expand = c(0, 0));rate_of_change

@@

@@

# Create box plot x axis as type (control and treatment) and y is HPI for kelp 
```{r}

# Plot three plots together using patchwork for production and HPI for fish and kelp biomass
```{r}
# Combine the plots with patchwork for all years of kelp and rate of change
production.hpi.nonlog <- production_plot | hpi.plot.nonlog +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi.nonlog
